---
title: "week16_recap"
author: "Daniel Trielli"
date: "2023-12-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## Loading the packages

Run the codeblock below to load the packages we will need for this recap

```{r}
library(tidyverse)
library(lubridate)
library(janitor)
library(tigris)
library(tidycensus)
library(corrr)
library(rvest)

library(tidytext)
library(tidyverse)
library(dplyr)
library(janitor)
library(lubridate)
library(rvest)
```

## Load Data

[USAFacts compiled data on inspections on child care facilities throughout the country](https://usafacts.org/data-projects/childcare-inspections) - specifically, how many of those facilities are overdue for inspection. Let's take a look on the data they have on Maryland.

First, here the data we need to load:

```{r}
childcare_inspections <- read_csv('data/childcare_inspections.csv')
childcare_inspections$fips <- as.character(childcare_inspections$fips)
```

#### **Q1** The childcare data does not have the names of the counties. Join them with the md_counties data, that can be loaded in the block below. Then inspect the dataframe you created with this join and answer this: there are two columns with county names in the new dataframe. Which one is the best to be used? (hint: it has to do with homonyms)

```{r}
# This loads the county data
md_counties <- counties() |>  filter(STATEFP == "24")

```

```{r}
# Join the childcare data with the county data 
counties_childcare <- left_join(childcare_inspections, md_counties, join_by("fips"== "GEOID"))
```

##### A1: The best column with county names to use is the NAMELSAD because Baltimore County and Baltimore City are both called Baltimore in the Name column so there would be no differentiation.

------------------------------------------------------------------------

#### **Q2** Create bar chart with the percentages of overdue childcare facility inspections per county to answer (You'll need to create a column for that calculation first). What is the county with the largest percentage of facilities with overdue inspections?

```{r}
# Calculate the percentage of overdue inspections per county
counties_childcare <- counties_childcare |>
  mutate(
    pct_overdue = (overdue/total)*100
  )
# Create the chart with ggplot
ggplot(counties_childcare, aes(x=NAMELSAD, y=pct_overdue)) +
  geom_bar(stat = "identity") 

```

##### **A2: The county with the largest percentage of facilities with overdue inspections is Montgomery County.**

------------------------------------------------------------------------

#### **Q3** Next, we want to see if the number of child care facilities - with overdue inspections or otherwise - is appropriate for each county. So we're going to calculate the rate of children per facility per county. To do that, you'll first need Census data to get the population of children under 5 for each county. Which county has the biggest rate of children per facility (i.e. they could use more facilities)?

```{r}
# First, grab data from the ACS. The ACS variable for children age under 5 is B06001_002 We want 2022 data.
census_api_key("7eeb7101193b9c08d0a867c4ea2547ecf86cd878", overwrite = TRUE)

md_under_5 <- get_acs(geography = "county",
              variables = c(underfive = "B06001_002"),
              state = "MD",
              year = 2022)
```

```{r}
# Rename the column name 'estimate', which has our population data, with something more useful, like 'children_pop'
md_under_5 <- md_under_5 |>
  rename('children_pop' = 'estimate')
```

```{r}
# Join the dataframes
age_counties_childcare <- md_under_5|> left_join(counties_childcare, by=c("GEOID" = "fips"))
# Calculate the rates for children per facility
age_counties_childcare <- age_counties_childcare |>
  mutate(
  rate = children_pop/total) |>
  arrange(desc(rate))
```

##### **A3: Garrett County has the biggest rate of children per facility with a rate of about 87.5 children per facility.**

------------------------------------------------------------------------

#### **Q4** Make a map of the results for question 3. Are there any areas of the state that are more lacking in childcare facilities?

```{r}
ggplot() +
  geom_sf(data = age_counties_childcare, aes(fill=rate, geometry=geometry)) +
  theme_minimal()
```

##### **A4**: Western Maryland , some of Central Maryland, and parts of the Eastern Shore seem to be lacking in childcare facilities more than the rest of the state.

------------------------------------------------------------------------

#### **Q5** Now let's explore if we can find patterns in under-served areas, particularly when it comes to median income of the population. First, join the childcare data median income data from the Census. Then, create a scatterplot and perform a cor.test(), and answer: is there is any correlation between income and rate of children per facilities? What makes you conclude that?

```{r}
# Again, grab data from the ACS. The ACS variable for median income is B19013_001. We want 2022 data again.
md_med_income <- get_acs(geography = "county",
              variables = c(medincome = "B19013_001"),
              state = "MD",
              year = 2022)
```

```{r}
# Rename the column name 'estimate', which has our income data, with something more useful, like 'median_income'
md_med_income <- md_med_income |> 
  rename('median_income' = 'estimate')
```

```{r}
# Join the dataframes
income_counties_childcare <- md_med_income |> left_join(age_counties_childcare, join_by("GEOID"))
```

```{r}
# Create a scatterplot with median income and rate of children per facility
ggplot(income_counties_childcare, aes(x=median_income, y=rate)) +
  geom_point()
```

```{r}
# Create a correlation test.
cor.test(income_counties_childcare$median_income, income_counties_childcare$rate)
```

##### A5: Given the scatter plot showing no correlation with the data points and a p value of 0.6067 (much greater than 0,05, meaning it's not statistically significant), we can see there is no correlation between median income and rate of children per facilities.

------------------------------------------------------------------------

#### **Q6** Finally, let's do some text analysis. We another dataset, with inspection reports from informal child care providers in the state. This dataset has an informative column of initial findings: a column that describes issues found in these providers, if any. Your goal is to create bigrams for this column and find common issues that are mentioned in these findings. And then answer: what kind of issues listed there you think are interesting? They don't have the most frequent.

```{r}
# Read the reports data
reports <- read_csv('data/childcare-informal-provider-inspections.csv') |> clean_names()
```

```{r}
# Create a stopword dictionary (feel free to include more words)
data("stop_words")
```

```{r}
# Unnest tokens into bigrams, removing stop words and repasting bigrams, list bigrams
reports |>
  unnest_tokens(bigram, initial_findings, token = "ngrams", n = 2) |>
  separate(bigram, c("word1", "word2"), sep = " ") |>
  filter(!word1 %in% stop_words$word) |>
  filter(!word2 %in% stop_words$word) |>
  mutate(bigram = paste(word1, word2, sep=" ")) |>
  group_by(bigram) |>
  tally(sort=TRUE) |>
  mutate(percent = (n/sum(n))*100) |>
  top_n(10)
```

##### A6: Some of the common issues listed that I think are interesting are emergency bags, cleaning products, missing items, smoke detector and outlet covers. I think these bigrams show some of the major issues and reasoning behind possible failed inspections of facilities.
