---
title: "Data Analysis Project"
author: "Marwa Barakat, Auzinea Bacon, Benji Ferraro"
date: "April 23"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this notebook, we are analyzing board game ranking data.

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(janitor)
library(lubridate)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data.

Source: Board Games Geek on Kaggle, the CSV was separated by semicolons instead of commas, so we used read_cvs2 instead of read_csv like normal. There are 14 columns of different aspects of each game and 20343 rows for each board game. As of initial analysis, there are some rows where a lot of the columns are a number that's not possible like 0, including year, play time and age. This is likely due to no collection of that data. The newest game is from 2022, so that limits us to that time and not using data or context from the past two years. We want to define complexity average because we have an assumption for what it might mean but will double check with original source. For the messed up years, we can verify the years on Google. This data doesn't include the brands/companies behind each game, which was something we were interested in looking into. For those with numbers/data that don't make sense, we will not put those into account when asking questions about those columns.

```{r}
board_game_data <- read_csv2("archive/bgg_dataset.csv") |>
  clean_names()

```

## Sentences to Engineer

In this notebook, we are analyzing five pieces from the data.

### Sentence 1

-   **Sentence text**: Compare the average rating of board games between 1900-2000 and 2000-2022.
-   **Analysis summary**:

```{r}
  

```

### Sentence 2

-   **Sentence text**: What age group is targeted most by these board games?
-   **Analysis summary**: In order to better look at age demographics, we grouped the minimum ages. Because the data only listed the minimum ages, we grouped them by 1-5, 6-12 , 13-17 and 18+. One important thing to note is that these aren't the age groups listed on each game but instead are the minimum ages for each game grouped together. For example, the 1-5 group is not games for ages 1 to 5. It is actually the games with minimum ages between 1 and 5 years old. The same applies to the other categories as well. After making these groups, we counted how many games fell under each one to see which age group is targeted most. There were over 14,000 games with minimum ages between 6 and 12 years old – more than any other age group. Because there were about 1,000 games with no minimum age listed, it was not possible for us to find the age of each board game, so it is important to note that those were not included in this analysis.

```{r}
board_game_ages <- board_game_data |>
  mutate(
    age_groups = case_when(
     between(min_age, 1, 5) ~ '1-5',
     between(min_age, 6, 12) ~ '6-12',
     between(min_age, 13, 17) ~ '13-17',
     min_age >= 18 ~ '18+',
     .default = 'N/A'
     )
  ) 

board_game_ages |>
  group_by(age_groups) |>
  summarise(
    total_games = n()
  ) |>
  filter(age_groups != 'N/A') |>
  arrange(desc(total_games))
```

### Sentence 3

-   **Sentence text**: Are solo board games more reviewed compared to multiplayer games?
-   **Analysis summary**:

```{r}
# This code shows the number of reviews for games that are considered multiplayer (2 or more players) compared to those that are considered singleplayer (no more than 1 player)

board_game_players <- board_game_data |>
  mutate(
    player_type = case_when(
    max_players <= 1~ "singleplayer",
    max_players >= 2 ~ "multiplayer",
    .default = NA
  ))

#Group by player type and count the number of ratings. Also, a sum of the number of games considered multiplayer and singleplayer for context.

board_game_players |> 
  group_by(player_type)|>
  summarise(
    total_ratings = sum(users_rated),
    total_games = n(),
    'avg_#_ratings_per_game' = total_ratings/total_games
    )
```

### Sentence 4

-   **Sentence text**: What years had the games owned by the most users?
-   **Analysis summary**:

```{r}


```

### Sentence 5

-   **Sentence text**:
-   **Analysis summary**:

```{r}


```

-30-
