---
title: "Data Analysis Project"
author: "Marwa Barakat, Auzinea Bacon, Benji Ferraro"
date: "April 23"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this notebook, we are reverse engineering the story, "[Inside the Political Donation History of Wealthy Sports Owners](https://fivethirtyeight.com/features/inside-the-political-donation-history-of-wealthy-sports-owners/)."

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}

library(tidyverse)
library(janitor)
library(lubridate)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data.

Source: Board Games Geek on Kaggle, the CSV was separated by semicolons instead of commas, so we used read_cvs2 instead of read_csv like normal. There are 14 columns of different aspects of each game and 20343 rows for each board game. As of initial analysis, there are some rows where a lot of the columns are a number that's not possible like 0, including year, play time and age. This is likely due to no collection of that data. The newest game is from 2022, so that limits us to that time and not using data or context from the past two years. We want to define complexity average because we have an assumption for what it might mean but will double check with original source. For the messed up years, we can verify the years on Google. This data doesn't include the brands/companies behind each game, which was something we were interested in looking into. For those with numbers/data that don't make sense, we will not put those into account when asking questions about those columns.

```{r}
board_game_data <- read_csv2("archive/bgg_dataset.csv") |>
  clean_names()


```

## Sentences to Engineer

In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

-   **Sentence text**:
-   **Analysis summary**:

```{r}

```

### Sentence 2

-   **Sentence text**:
-   **Analysis summary**:

```{r}


```

### Sentence 3

-   **Sentence text**:
-   **Analysis summary**:

```{r}


```

### Sentence 4

-   **Sentence text**:
-   **Analysis summary**:

```{r}


```

### Sentence 5

-   **Sentence text**:
-   **Analysis summary**:

```{r}


```

-30-
