---
title: "Data Analysis Project"
author: "Marwa Barakat, Auzinea Bacon, Benji Ferraro"
date: "April 23"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this notebook, we are analyzing board game ranking data.

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(janitor)
library(lubridate)

options(scipen = 999)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data.

Source: Board Games Geek on Kaggle, the CSV was separated by semicolons instead of commas, so we used read_cvs2 instead of read_csv like normal. There are 14 columns of different aspects of each game and 20343 rows for each board game. As of initial analysis, there are some rows where a lot of the columns are a number that's not possible like 0, including year, play time and age. This is likely due to no collection of that data. The newest game is from 2022, so that limits us to that time and not using data or context from the past two years. We want to define complexity average because we have an assumption for what it might mean but will double check with original source. For the messed up years, we can verify the years on Google. This data doesn't include the brands/companies behind each game, which was something we were interested in looking into. For those with numbers/data that don't make sense, we will not put those into account when asking questions about those columns.

```{r}
board_game_data <- read_csv2("archive/bgg_dataset.csv") |>
  clean_names()

```

## Analysis Questions

In this notebook, we are analyzing/answering five questions using this board game data set.

### Question 1

-   **Question**: Compare the average rating of board games between 1900-2000 and 2000-2022.

-   **Analysis summary**:

    We mutated the data so that we could compare both centuries in terms of average rating for board games. We mutated the years published and grouped by time span for average rating, and got approximately 6.5 for the average rating from 2000 to 2022, and 6 for the average rating from 1900-1999. In terms of modern board games in this data set, it seems that the 21st century has a higher average score. This could be due to differing total amounts for each of the centuries, which could skew the data.

```{r}
board_game_years <- board_game_data |>
  mutate(
    era = case_when(
     between(year_published, 1900, 1999) ~ '1900-1999',
     between(year_published, 2000, 2022) ~ '2000-2022',
     .default = 'N/A'
     ))

board_game_years |>
  group_by(era) |>
  summarise(
    avg_rating = mean(rating_average)
  ) |>
  filter(era != 'N/A') |>
  arrange(desc(avg_rating))
```

### Question 2

-   **Question**: What age group is targeted most by these board games?
-   **Analysis summary**: In order to better look at age demographics, we grouped the minimum ages. Because the data only listed the minimum ages, we grouped them by 1-5, 6-12 , 13-17 and 18+. One important thing to note is that these aren't the age groups listed on each game but instead are the minimum ages for each game grouped together. For example, the 1-5 group is not games for ages 1 to 5. It is actually the games with minimum ages between 1 and 5 years old. The same applies to the other categories as well. After making these groups, we counted how many games fell under each one to see which age group is targeted most. There were over 14,000 games in this data set with minimum ages between 6 and 12 years old – more than any other age group. Because there were about 1,000 games with no minimum age listed, it was not possible for us to find the age of each board game, so it is important to note that those were not included in this analysis.

```{r}
board_game_ages <- board_game_data |>
  mutate(
    age_groups = case_when(
     between(min_age, 1, 5) ~ '1-5',
     between(min_age, 6, 12) ~ '6-12',
     between(min_age, 13, 17) ~ '13-17',
     min_age >= 18 ~ '18+',
     .default = 'N/A'
     )
  ) 

board_game_ages |>
  group_by(age_groups) |>
  summarise(
    total_games = n()
  ) |>
  filter(age_groups != 'N/A') |>
  arrange(desc(total_games))
```

### Question 3

-   **Question**: Are solo board games more reviewed compared to multiplayer games?

-   **Analysis summary**:

    Single player games have fewer reviews than multi player games in this data set, as shown by the total ratings column below. There are about 168,000 more reviews for multi player games than single player games. However, there is also a much larger number of multi player games overall when compared to single player - there are less than 500 single player games in the review list. This means that there are statistically more opportunities to review multi player games than single player and, potentially, groups involved in playing multi player games may individually leave multiple reviews for those games, resulting in many more ratings for those games than single players. We calculated an average number of ratings to account for this difference, showing multi player games in this data set on average have a higher number of ratings.

```{r}
# This code shows the number of reviews for games that are considered multiplayer (2 or more players) compared to those that are considered singleplayer (no more than 1 player)



board_game_players <- board_game_data |>
  mutate(total_players = min_players+max_players) |>
  mutate(
    player_type = case_when(
    total_players <= 1~ "singleplayer",
    total_players >= 2 ~ "multiplayer",
    .default = NA
  ))


#Group by player type and count the number of ratings. Also, a sum of the number of games considered multiplayer and singleplayer for context.


board_game_player_ages <- board_game_players |> 
  group_by(player_type)|>
  summarise(
    total_ratings = sum(users_rated),
    total_games = n()
    )

board_game_player_ages |>
  mutate(
    avg_ratings = (total_ratings/sum(total_ratings))*100
  ) |>
  arrange(desc(avg_ratings))

```

### Question 4

-   **Question**: What decade had games owned by the most users?
-   **Analysis summary**: To answer this question, we first made a new column grouping together games published in the same decade starting from the year 1900. Then, we made a table showing how many total games fall under each decade, how many total owners there are of games in that decade, and the average number of owners per game in each decade. Then, to further learn more from the data, we calculated the average number of owners in each decade by comparing it to the total owners for all of those decades. This way the averages can be better compared to each other. We found that games published between 2010 and 2019 in this data set had the highest average owners compared to the other decades.

```{r}

board_game_owners <- board_game_data |>
   mutate(
    decades = case_when(
     between(year_published, 1900, 1909) ~ '1900-1909',
     between(year_published, 1910, 1919) ~ '1910-1919',
     between(year_published, 1920, 1929) ~ '1920-1929',
     between(year_published, 1930, 1939) ~ '1930-1939',
     between(year_published, 1940, 1949) ~ '1940-1949',
     between(year_published, 1950, 1959) ~ '1950-1959',
     between(year_published, 1960, 1969) ~ '1960-1969',
     between(year_published, 1970, 1979) ~ '1970-1979',
     between(year_published, 1980, 1989) ~ '1980-1989',
     between(year_published, 1990, 1999) ~ '1990-1999',
     between(year_published, 2000, 2009) ~ '2000-2009',
     between(year_published, 2010, 2019) ~ '2010-2019',
     between(year_published, 2020, 2022) ~ '2020-2022',
     .default = 'N/A'
     ))

board_game_owner_decades <- board_game_owners |>
  group_by(decades) |>
  filter(owned_users != 'N/A') |>
  filter(decades != 'N/A') |>
  summarise(
    total_games = n(),
    total_owners = sum(owned_users)
    ) 

board_game_owner_decades |>
  mutate(
    avg_per_game = total_owners/total_games,
    avg_owners = (total_owners/sum(total_owners))
  ) |>
  arrange(desc(avg_owners))
```

### Question 5

-   **Question**: Compare two of the most common domains and how many games there are of them.
-   **Analysis summary**: First, we wanted to see which domains were the most popular that the baord games in this data fell under. After analyzing the data and noticing that strategy games and wargames were the most common domains, we decided to compare those two categories alone. Since the domain category listed all applicable categories, we first used strdetect to find all of the games with "Strategy Games" as one of their domains and counted how many there were. Next, we did the same for "Wargames." After doing this, we found that this data set had 3316 wargames and 2205 strategy games, showing that from the games in this data set, wargames are more common.

```{r}
#This is to see which domains are most common in this data set
board_game_domains <- board_game_data |>
  group_by(domains) |>
  summarise(
    count = n()
  ) |>
  arrange(desc(count))


strategy <- board_game_data |>
  filter(str_detect(domains, "Strategy Games")) |>
  group_by(domains) |>
  summarise(
    total_games = n(),
    ) |>
  mutate(
    total = sum(total_games)
  )



war <- board_game_data |>
  filter(str_detect(domains, "Wargames")) |>
  group_by(domains) |>
  summarize(
    total_games = n()) |>
  mutate(
    total = sum(total_games)
  )
  

```

-30-
