---
title: "Reverse Engineering Project"
author: "Marwa Barakat, Laura Van Pate, Judith Altneu"
date: "March 27"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this notebook, we are reverse engineering the story, "[Inside the Political Donation History of Wealthy Sports Owners](https://fivethirtyeight.com/features/inside-the-political-donation-history-of-wealthy-sports-owners/)"

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}

library(tidyverse)
library(janitor)
library(lubridate)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data.

```{r}
# Load required data

donation_data <- read_csv('data/sports-political-donations.csv') |>

clean_names()

party_donations <- donation_data |>
  mutate(
    new_party = case_when(
      party == 'Republican' ~ 'Republican',
      party == 'Democrat' ~ 'Democrat' ,
      .default = 'Bipartisan'
    )
  ) |>
  mutate(new_amount = parse_number(amount))

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 

```

## Sentences to Engineer

In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

-   **Sentence text**: "this deep-pocketed group has sent \$34.2 million (72.9 percent) to Republican campaigns or super PACs purely supporting Republican causes, compared to \$10.1 million (21.5 percent) to Democrats over the past three elections."
-   **Analysis summary**: I was able to confirm the finding that sports owners have sent \$34.2 million to Republicans and \$10.1 million to Democrats over the past three elections. I was also able to confirm the percentages of 72.9 and 21.5 through my analysis of the data.

```{r}
# Put code to reverse engineer sentence here


party_donations |>
 group_by(new_party) |>
  summarise(
    total_amount = sum(new_amount)
    )|>
  arrange(desc(total_amount)) |>
  mutate(
    total_donations = sum(total_amount),
    pct_total = total_amount/total_donations
  )

  
# Display results of code below this codeblock
```

### Sentence 2

-   **Sentence text**: there are 40 owners who have donated at least \$100,000 to Republican causes over the 6 year period
-   **Analysis summary**: I was able to confirm this finding. At first, I thought the data was missing something because the table came out to have only 37 owners but I realized this was due to some of the owners donating together (like wife and husband), so if you count each individual, there are 40 total.

```{r}
# Put code to reverse engineer sentence here

republican_donations <- party_donations |>
  filter(str_detect(party, 'Republican')) |>
  group_by(owner) |>
  summarise(
    total_donations = sum(new_amount)
  ) |>
  filter(total_donations >= 100000) |>
  arrange(desc(total_donations))

# Display results of code below this codeblock

```

### Sentence 3

-   **Sentence text**: 23 owners donated over \$100,000 to Democrat causes and candidates
-   **Analysis summary**: I was able to confirm this finding. Similarly, I thought something was wrong in the data because it came out to 22 not 23, but I realized this was due to a couple donating together, so if you count each individual, there are 23 total.

```{r}
# Put code to reverse engineer sentence here

democrat_donations <- party_donations |>
  filter(str_detect(party, 'Democrat')) |>
  group_by(owner) |>
  summarise(
    total_donations = sum(new_amount)
  ) |>
  filter(total_donations >= 100000) |>
  arrange(desc(total_donations))

# Display results of code below this codeblock

```

### Sentence 4

-   **Sentence text**: MLB owners have donated the most (the table)
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

party_donations |> 
  group_by(league) |>
  summarise(
    total_democratic = sum(new_amount[new_party == 'Democrat']),
    total_republican = sum(new_amount[new_party == 'Republican']),
    total_donations = sum(new_amount)
  ) |>
  arrange(desc(total_donations))

# Display results of code below this codeblock

```

### Sentence 5

-   **Sentence text**: Thirteen owners have contributed directly to Trump's campaign or Trump super PACs in the 2020 election cycle, with contributions adding up to \$1,746,423
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-
